---
title: "Data exploration for 'survey' dataset"
author: "Henry Baker"
date: "2023-12-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readr)
library(dplyr)
library(haven)
library(plotly)
library(patchwork)
library(tidyr)
library(fuzzyjoin)
```

## Loading & initial look at data

```{r survey data folder}

setwd("/Users/henrybaker/Library/Mobile Documents/com~apple~CloudDocs/Documents/GitHub/IDS'23/hackathon-project-individual-henrycgbaker")

survey_data <- readRDS("survey/meof_usa_survey_df.RDS")

#str(survey_data)
#summary(survey_data)

survey_data_variables <- names(survey_data)

```

```{r tracking data folder}
#tracking
tracking_data <- readRDS("tracking/meof_usa_web_df.RDS")
#str(tracking_data)
#summary(tracking_data)

tracking_data_variables <- names(tracking_data)

# time zones
time_zones <- readRDS("tracking/time_zones.RDS")
#str(time_zones)
#summary(time_zones)
```

## Attempted to do a data mining attempt here (failed)
this was meant to look at the covariance between all columns to see if any initial patterns & correlations emerged, but the data isn't properly numeric and most of it drops out.
```{r covariance between columns}
# doesn't work - too many omitted columns:

selected_data <- select(survey_data, -c(1:7))
# Identify numeric and convertible columns
numeric_columns <- sapply(selected_data, function(x) is.numeric(x) || is.factor(x))

# Convert  to numeric
survey_numeric <- data.frame(lapply(selected_data[, numeric_columns, drop = FALSE], function(x) {
  if (is.factor(x)) {
    as.numeric(as.character(x))
  } else {
    as.numeric(x)
  }
}))

#  compute covariance
vc_survey <- cov(survey_numeric)

```

## basic transfomrations to make data more useable moving forwards
Continue to add to this as make basic changes - group them altogether here so can keep track
```{r basic repub demographics}

#create new party variable for ease of use
survey_data$party <- ifelse(survey_data$pid3 == 1, "democrat", "republican")

# create a  republican subset
repub_subset <- survey_data |>
  filter(party == "republican") |>
  group_by(personid)

n_repub <- nrow(repub_subset)

# create a democratic subset
demo_subset <- survey_data |>
  filter(party == "democrat") |>
  group_by(personid)

n_demo <- nrow(demo_subset)
# NB signifanctly less democrats !
```

## Basic demographic distributions

### Republicans
```{r}

par(mfrow = c(3, 3))

# Gender
gender_plot <- ggplot(repub_subset, aes(x = factor(gender))) +
  geom_bar() +
  labs(title = "Gender Distribution among Republican Respondents", x = "Gender", y = "Count")

# Birth Year
birthyr_plot <- ggplot(repub_subset, aes(x = birthyr)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Birth Year Distribution among Republican Respondents", x = "Birth Year", y = "Count")

# Race
race_plot <- ggplot(repub_subset, aes(x = factor(race))) +
  geom_bar() +
  labs(title = "Race Distribution among Republican Respondents", x = "Race", y = "Count")

# Education
educ_plot <- ggplot(repub_subset, aes(x = factor(educ))) +
  geom_bar() +
  labs(title = "Education Level Distribution among Republican Respondents", x = "Education Level", y = "Count")

# Family Income
faminc_plot <- ggplot(repub_subset, aes(x = factor(faminc_new))) +
  geom_bar() +
  labs(title = "Family Income Distribution among Republican Respondents", x = "Income Level", y = "Count")

# Religious Pew
regligpew_plot <- ggplot(repub_subset, aes(x = factor(religpew))) +
  geom_bar() +
  labs(title = "Religious Affiliation Distribution among Republican Respondents", x = "Religious Affiliation", y = "Count")

# Combine the plots in a 3x2 layout
n_label_plot_repub <- ggplot() +
  annotate("text", x = 0, y = 0, label = paste("Total n =", n_repub), size = 5) +
  theme_void()

combined_plot <- gender_plot / birthyr_plot / race_plot / educ_plot / faminc_plot / regligpew_plot / n_label_plot_repub + 
                 plot_layout(ncol = 2, nrow = 4)

combined_plot

```

For religious affiliation: 
1 = prot
2 = catholic
11 = nothing in particular

### Democrats

```{r basic democrat demographics}

par(mfrow = c(3, 2))

# Gender
gender_plot <- ggplot(demo_subset, aes(x = factor(gender))) +
  geom_bar() +
  labs(title = "Gender Distribution among Democratic Respondents", x = "Gender", y = "Count")

# Birth Year
birthyr_plot <- ggplot(demo_subset, aes(x = birthyr)) +
  geom_histogram(binwidth = 10) +
  labs(title = "Birth Year Distribution among Democratic Respondents", x = "Birth Year", y = "Count")

# Race
race_plot <- ggplot(demo_subset, aes(x = factor(race))) +
  geom_bar() +
  labs(title = "Race Distribution among Democratic Respondents", x = "Race", y = "Count")

# Education
educ_plot <- ggplot(demo_subset, aes(x = factor(educ))) +
  geom_bar() +
  labs(title = "Education Level Distribution among Democratic Respondents", x = "Education Level", y = "Count")

# Family Income
faminc_plot <- ggplot(demo_subset, aes(x = factor(faminc_new))) +
  geom_bar() +
  labs(title = "Family Income Distribution among Democratic Respondents", x = "Income Level", y = "Count")

# Religious Pew
regligpew_plot <- ggplot(demo_subset, aes(x = factor(religpew))) +
  geom_bar() +
  labs(title = "Religious Affiliation Distribution among Republican Respondents", x = "Religious Affiliation", y = "Count")

# Combine the plots in a 3x2 layout
n_label_plot_demo <- ggplot() +
  annotate("text", x = 0, y = 0, label = paste("Total n =", n_demo), size = 5) +
  theme_void()

combined_plot <- gender_plot / birthyr_plot / race_plot / educ_plot / faminc_plot / regligpew_plot / n_label_plot_demo + 
                 plot_layout(ncol = 2, nrow = 4)

combined_plot

```
NB: many fewer democrats in the sample

## System trust by party

```{r system trust}

ggplot(survey_data, aes(x = system_fair, fill = party)) +
  geom_histogram(position = "dodge", binwidth = 1) +
  scale_fill_manual(values = c("democrat" = "blue", "republican" = "red")) +
  labs(title = "Distribution of System Fairness Perception by Party",
       x = "System Fairness Rating",
       y = "Count") +
  theme_minimal()

```

1 = strongly disagree
5 = strongly agree

this is interesting - more antisystem sentiment amongst democrats
Also high variance - definitely an interesting variable to unpack
- what predictors for system_fair in survey data
- would be very interesting to see how tracking data maps on to this
- changes over time? recorded in all 7 waves: look how system trust changes for specific subsets within parties over time?
- does it change with changes in tracking data website visits and news consumption
- is it associated with non legacy media consumption? (age interaction term?) 
- diff in diff model: get baseline of similar units, then see how consumption of certain media in intervention group changes trust in system compared to control units?
- do late night users have less trust?

## Platform statistics by party

```{r platform stats}

# function to create plot for a social media platform
platform_plot_fn <- function(data, platform, party_col = "party") {
  ggplot(data, aes_string(x = party_col, fill = party_col)) +
    geom_bar(data = subset(data, get(platform) == 1)) +
    labs(title = paste("Users with", platform, "Accounts by Party"),
         x = "Party",
         y = "Count") +
    scale_fill_manual(values = c("democrat" = "blue", "republican" = "red")) +
    theme_minimal()
}
  
# iterate
twitter_plot <- platform_plot_fn(survey_data, "accounts_twitter")
facebook_plot <- platform_plot_fn(survey_data, "accounts_facebook")
instagram_plot <- platform_plot_fn(survey_data, "accounts_instagram")
linkedin_plot <- platform_plot_fn(survey_data, "accounts_linkedin")

combined_plot <- twitter_plot / facebook_plot / instagram_plot / linkedin_plot +
                 plot_layout(ncol = 2)

combined_plot

```

but this is because there are more republicans
need to work out as proportion

```{r proportional platform stats}

# function to calc props for each platform
calculate_proportions <- function(data, platform) {
  data |>
    group_by(party) |>
    summarise(Proportion = mean(get(platform) == 1, na.rm = TRUE), .groups = 'drop')
}

# calc propsfor each platform
twitter_proportions <- calculate_proportions(survey_data, "accounts_twitter")
facebook_proportions <- calculate_proportions(survey_data, "accounts_facebook")
instagram_proportions <- calculate_proportions(survey_data, "accounts_instagram")
linkedin_proportions <- calculate_proportions(survey_data, "accounts_linkedin")

twitter_proportions
facebook_proportions
instagram_proportions
linkedin_proportions

# into one data frame
combined_proportions <- bind_rows(
  mutate(twitter_proportions, Platform = "Twitter"),
  mutate(facebook_proportions, Platform = "Facebook"),
  mutate(instagram_proportions, Platform = "Instagram"),
  mutate(linkedin_proportions, Platform = "LinkedIn")
)

# long format for plots
long_proportions <- pivot_longer(combined_proportions, 
                                 cols = -c(party, Platform), 
                                 names_to = "SocialMedia", 
                                 values_to = "Proportion")

ggplot(long_proportions, aes(x = party, y = Proportion, fill = party)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  facet_wrap(~ Platform) +
  labs(title = "Proportion of Each Party's Members Using Social Media Platforms",
       x = "Party",
       y = "Proportion") +
  scale_fill_manual(values = c("democrat" = "blue", "republican" = "red")) +
  theme_minimal()

```


## to do:
- once have more specific set of variables, drop unnecessary then join the two data sets together in advance - there's only one way (join by personid) and we will def have to do this

